A meta-analysis of fungicide spray schedules to minimise yield loss from powdery mildew in Australian mungbean

### 2020-08-10 02_Import_and_FormatData.Rmd
rename files and delete unnecessary files/chapters

```{r PA_Libraries, include=FALSE, message=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               lubridate,
               bomrang,
               openxlsx,
               devtools,
               reshape2,
               lme4,
               kableExtra)
source("R/import_data.R")
if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq()
)
```

>>>>

```{r PA_Libraries, include=FALSE, message=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               lubridate,
               bomrang,
               openxlsx,
               devtools,
               reshape2,
               lme4,
               kableExtra,
               here)
source(here("R/import_data.R"))
if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq()
)
```

### 2020-08-10 04_Prep_data.Rmd
rename files and delete unnecessary files/chapters

@@

```{r load_libraries}
library(dplyr)
library(readr)
library(kableExtra)
library(ggplot2)
library(agricolae)
```

```{r import_data}
source("R/import_data.R")
PM_MB_means <- import_data()

# Calculate AUDPC using final severity and first sign of disease - 7 days
source("R/AUDPC_for_single_obs.R") 
```

@@

We are narrowing the meta-analysis to only include demethylation inhibitors, tebuconazole and propiconazole. Within studies which have used these fungicides, what type of error was reported?

```{r variance_types}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole",
         row_spacing == 0.75) %>%
   group_by(trial_ref, location, Y_error_type, fungicide_ai) %>%
   summarise(DMI_trts = length(Y_error_type))%>%
   arrange(Y_error_type)%>%
   kable(caption = "Number of treatments for each trial and the error type reported") %>%
   kable_styling(fixed_thead = T) %>%
   footnote(general = "DMI - demethylation inhibitors (tebuconazole and propiconazole).")

```

The experiment summaries report either standard deviation, least square differences or no error term (`NA`). Below we will convert LSD and standard deviation to sample variance. In addition, we need to impute variances for the trials which did not report any form of variance (`NA`).

@@

```{r all_LSD_trials}
Trials <- PM_MB_means %>%
   filter(
          Y_error_type == "lsd (P=0.05)") %>%
   select(trial_ref) %>%
   distinct()
```

The following code attempts to convert 'least square differences' (LSD) to sampling variance as per the method in [Nugugi et.al (2011)](https://apsjournals.apsnet.org/doi/10.1094/PHYTO-08-10-0221).
Ideas for using a T-critical value of 1.697 came from reading a [statisics-how-to website](https://www.statisticshowto.datasciencecentral.com/how-to-calculate-the-least-significant-difference-lsd/).
A [table of T-critical values](https://www.statisticshowto.datasciencecentral.com/tables/t-distribution-table/) was consulted where 30 degrees of freedom within experiment was used to find the 'T-crit' value of 1.697.

@@

```{r LSD_2_Sampling Variance}

for (Trial2012 in Trials$trial_ref) {
   Tcrit <-
      1.697  # 0.05 and DFw = 30 # T critical value for which there is significant difference between two groups; This value is the same for both trials "mung1112/01" and "mung1112/02"
   DFw <-
      sum(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates) -
      length(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates) # degrees of freedom within groups

   LSD <-
      PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$yield_error[1]

   V_yield <-
      (PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[1] * (LSD /
                                                                            Tcrit) ^ 2) / 2 # method - Nugugi et.al (2011) DOI: 10.1094/phyto-08-10-0221

   for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates)) {
      if (j == 1) {
         Sv <-
            vector(length = length(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates))
      } # reset Sv vector if this is the first loop

      ###############################################3
      # ___ Adjusting formula to suit our data_______
      #
      #  Sv[j]^2 = MSE/(PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$replicates[j] *
      #                   PM_MB_means[PM_MB_means$trial_ref == Trial2012,]$grain_yield.t.ha[j])
      # Balance equation
      # Sv[j] = V_yield / 
      #    (PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[j] *
      #       as.numeric(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$grain_yield.t.ha[j])^2)
      # #################################################################

      Sv[j] = V_yield / 
         PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[j]

      # Sv[j] = MSE / 
      #    (PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$replicates[j] *
      #       as.numeric(PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$grain_yield.t.ha[j])^2)
      #

      ### -> This formula should be V = (n * (LSD/1.96)^2) / 2

@@

   # and replacing the values in the dataset

   PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$yield_error <-
      Sv
   PM_MB_means[PM_MB_means$trial_ref == Trial2012, ]$Y_error_type <-
      "SamplingVariance"

}
```

### Converting standard deviation to sample variance
Now to estimate the variance from the standard deviation

@@

# and replacing the values in the dataset
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error <-
   Sv
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$Y_error_type <-
   "SamplingVariance"

```

## Yield saving  
Yield saving and Yield saving proportional to the control.

Data from `r length(PM_MB_means$trial_ref)` trials were collected and the means of each treatment calculated (`R/Cleaning historical data.R`). These data were combined into one file `1902 powdery mildew-Mungbean - Collated means.csv`.

```{r Calculate_yield-gains}
PM_MB_means$grain_yield.t.ha. <-
   as.numeric(PM_MB_means$grain_yield.t.ha.) # NAs produced due to some cells having text description to why there is no specific data

for (i in unique(PM_MB_means$trial_ref)) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]

   if (any(is.na(unique(dat1$row_spacing)))) {
      warning(
         paste(
            unique(dat1$trial_ref),
            "at",
            unique(dat1$location),
            "in year",
            unique(dat1$year),
            "has unknown row.spacing (NA)\n 'yield_gain' and 'proportional yield gain' not calculated\n"
         )
      )
      next()
   }

   for (j in unique(dat1$row_spacing)) {
      dat2 <- dat1[dat1$row_spacing == j, ]
      controlY <-
         mean(dat2[dat2$fungicide_ai == "control", "grain_yield.t.ha."])

      dat2$yield_gain <- dat2$grain_yield.t.ha. - controlY
      dat2$prop_YG <- dat2$yield_gain  /  controlY

      dat2[dat2$fungicide_ai == "control", c("yield_gain", "prop_YG")] <-
         NA

      dat1[dat1$row_spacing == j, ] <- dat2
   }
   PM_MB_means[PM_MB_means$trial_ref == i, ] <- dat1
}

# ____________________________________________
# calculate yield gain and prop_yield gain for experiments with no row spacing data

dat1 <- PM_MB_means[PM_MB_means$trial_ref == "AM1303", ]

controlY <-
   mean(dat1[dat1$fungicide_ai == "control", "grain_yield.t.ha."])

dat1$yield_gain <- dat1$grain_yield.t.ha. - controlY
dat1$prop_YG <- dat1$yield_gain / controlY
dat1[dat1$fungicide_ai == "control", c("yield_gain", "prop_YG")] <- c(NA,NA)
PM_MB_means[PM_MB_means$trial_ref == "AM1303", ] <- dat1

write.csv(PM_MB_means,
          "data/1911_PM_MB_means&Ygains.csv")
```

## Reduce data set to required variables  
Lets tidy up the data set into a data frame of only the variables which are required for the analysis.

First remove the irrelevant columns/variables. Then calculate fungicide application timing variables relative to the first sign of disease. Entries with NA in the grain yield or powdery mildew severity are also removed from the data.

```{r clustered_fungicide_applications}
source("R/slimming_PM_dat.R") # see R script for the code which reduces the data.

slim_PM_dat %<>%
  mutate(fungicide_timing_1 = fungicide_application_1 - first_sign_disease) %>%
  mutate(fungicide_timing_2 = fungicide_application_2 - fungicide_application_1) %>%
  mutate(fungicide_timing_3 = fungicide_application_3 - fungicide_application_2) %>%
  filter(!is.na(grain_yield.t.ha)) %>% 
  filter(!is.na(PM_final_severity))

```

### Cluster fungicide applications  
We are going to define 'fungicide application timing' as a categorical variable, relative to the first sign of powdery mildew.

These categorical variables are named:  
   - **Early**: First fungicide application was prior to first sign of disease.  
   - **Recommended**: First fungicide application was applied on the day powdery mildew was observed, or within three days of first sign.  
   - **Late**: First fungicide application was four or more days after first sign of disease being observed.

Follow up fungicide applications (cluster_2 and cluster_3) were also categorised as either, early, recommended or late.

```{r cluster}

slim_PM_dat %<>%
  mutate(
    cluster_1 = case_when(
      fungicide_timing_1 < 0 ~ "Early",
      fungicide_timing_1 >= 0 &
        fungicide_timing_1 < 4 ~ "Recommended",
      TRUE ~ "Late"
    )
  ) %>%
  mutate(
    cluster_2 = case_when(
      is.na(fungicide_timing_2) ~ NA_character_,
      fungicide_timing_2 < 13 ~ "Early",
      fungicide_timing_2 >= 13 &
        fungicide_timing_2 < 17 ~ "Recommended",
      TRUE ~ "Late"
    )
  ) %>% 
  mutate(
    cluster_3 = case_when(
      fungicide_timing_3 < 14 ~ "Early",
      is.na(fungicide_timing_3) ~ NA_character_,
      TRUE ~ "Recommended"
    )
  )

```

Lets look at these clusters to see where we have the most degrees of freedom.

```{r cluster_clusters}

clusters <- unique(slim_PM_dat[c("cluster_1","cluster_2","cluster_3")])

for(i in 1:nrow(clusters)){
   if(i == 1){clusters$n <- NA
   clusters$n_x_trial <- NA}

   clusters[i,"n"] <- nrow(slim_PM_dat[slim_PM_dat$cluster_1 == clusters[i,"cluster_1"] &
                                          slim_PM_dat$cluster_2 == clusters[i,"cluster_2"] &
                                          slim_PM_dat$cluster_3 == clusters[i,"cluster_3"],
                                       ])

   clusters[i,"n_x_trial"] <- length(unique(slim_PM_dat[slim_PM_dat$cluster_1 == clusters[i,"cluster_1"] &
                                          slim_PM_dat$cluster_2 == clusters[i,"cluster_2"] &
                                          slim_PM_dat$cluster_3 == clusters[i,"cluster_3"],
                                       "trial_ref"]))

@@

knitr::kable(clusters)

```

From this we can see that there are 16 different cluster combinations, which would make the analysis too difficult. I will reduce this to 6 levels within one variable.

Single sprays either early, recommended timing or late.

Two sprays or more where the first spray was either early, recommended timing, or late.

```{r simple_clusters}

slim_PM_dat <- slim_PM_dat %>%
   mutate(
    spray_management = case_when(
       fungicide_timing_1 < 0 &
          is.na(fungicide_application_2) &
          is.na(fungicide_application_3) ~ "Early",
       fungicide_timing_1 >= 0 &
          fungicide_timing_1 < 4 &
          is.na(fungicide_application_2) &
          is.na(fungicide_application_3) ~ "Recommended",
      fungicide_timing_1 >= 4 &
          is.na(fungicide_application_2) &
          is.na(fungicide_application_3) ~ "Late",
      fungicide_timing_1 < 0 &
          !is.na(fungicide_application_2) ~ "Early_plus",
       fungicide_timing_1 >= 0 &
          fungicide_timing_1 < 4 &
          !is.na(fungicide_application_2) ~ "Recommended_plus",
      fungicide_timing_1 >= 4 ~ "Late_plus",
      TRUE ~ "Other"
    )
  )

slim_PM_dat[slim_PM_dat$fungicide_ai == "control",
            c("fungicide_timing_1", "fungicide_timing_2", "fungicide_timing_3",
              "spray_management"
              )] <- "control"

table(slim_PM_dat$spray_management)

table(slim_PM_dat$spray_management,slim_PM_dat$fungicide_timing_1)

```
There are few 'Early_plus' and 'late' treatments, these treatments therefore will have too few comparisons with other treatments in the meta-analysis to be provide accurate results. Therefore we will remove 'Early_plus' and 'Late' from the analysis.

```{r simpler_clusters_REMOVE}

slimmer_PM_dat <- slim_PM_dat[slim_PM_dat$spray_management != "Late" & 
               slim_PM_dat$spray_management != "Early_plus",]

write.csv(slimmer_PM_dat, file = "data/slimmer_PM_clusterdat.csv", row.names = FALSE)
```

Our data currently uses sample variance to define the yield error. We should expect that when yield increases so does the variance. Lets quickly plot this to check.

```{r Var_mean_sqaures}
# redefine sample variance calculated from mean square
slimmer_PM_dat$vi <- 
   slimmer_PM_dat$Y_Msquare / 
   (slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha ^ 2)

slimmer_PM_dat %>%
   ggplot(aes(x= grain_yield.t.ha, y = vi))+
   geom_point()
```

We see here that the sample variance generally decreases or has no association with the grain yield. This might be due to the unpredictable nature of mungbeans, however going forward we will use mean squares which defines between trial variation and is a the more generic error term.

```{r Var_MeanSquarePlot}
slimmer_PM_dat %>%
   ggplot(aes(x= grain_yield.t.ha, y = Y_Msquare))+
   geom_point()

```

```{r Yield_vs_AUDPC}

slimmer_PM_dat %>%
   filter(AUDPC_m <=280)%>%
#   filter(fungicide_ai == "control")%>%
   ggplot(aes(x = grain_yield.t.ha, y = AUDPC_m))+
   geom_point()

summary(lm(grain_yield.t.ha ~ AUDPC_m, data = subset(slimmer_PM_dat, AUDPC_m <= 280)))

plot(grain_yield.t.ha. ~ AUDPC_m,data = subset(PM_MB_means, AUDPC_m <= 280) )
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = subset(PM_MB_means, AUDPC_m <= 280)))

sm2lg <- PM_MB_means%>%
   group_by(trial_ref)%>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE))%>%
   arrange(TY)

for(i in sm2lg$trial_ref){
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]
   if(any(is.na(dat1$grain_yield.t.ha.))){next}
   if(any(is.na(dat1$AUDPC_m))){next}

plot(grain_yield.t.ha. ~ AUDPC_m, 
     data = dat1, main = paste(dat1$location[1],(dat1$year[1])),
     ylim = c(0,max(PM_MB_means$grain_yield.t.ha., na.rm = TRUE)), 
     xlim = c(0,max(PM_MB_means$AUDPC_m, na.rm = TRUE)))
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = dat1))
points(grain_yield.t.ha. ~ AUDPC_m, data = dat1[dat1$fungicide_ai == "control",], col = "red")

```

This sequence of graphs show that AUDPC and disease incidence does not impact on yields if the season is a low yielding, ie lower than 1 tonne per hectare.  
The higher the yield the greater greater the influence on the disease on the yield especially if the crop is high yielding and a large AUDPC.

```{r delete_this_chunk}
PM_MB_means$infected_days <- as.numeric(PM_MB_means$harvest_date - PM_MB_means$first_sign_disease)

plot(AUDPC_m ~ infected_days, data = PM_MB_means)
plot(grain_yield.t.ha. ~ infected_days, data = PM_MB_means)

PM_MB_means$harvest_date - PM_MB_means$first_sign_disease
PM_MB_means[0 > as.Date(PM_MB_means$harvest_date, format = "%Y-%m-%d")-
as.Date(PM_MB_means$first_sign_disease, format = "%Y-%m-%d"),]

# correct dates (this has been done in wrangle file but not updated yet in the data-set)
PM_MB_means[PM_MB_means$trial_ref == "mung1516/01", "harvest_date"] <- as.Date("2016-05-06", format = "%Y-%m-%d")
PM_MB_means[PM_MB_means$trial_ref == "mung1516/01", "final_assessment"] <- final_assessment = as.Date("2016-04-15", format = "%Y-%m-%d")

#curiouspw10-
```

>>>>

# Prepare data for meta-analysis

@@

```{r load_libraries, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, kableExtra, agricolae, lme4, here)
if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq()
)
```

```{r import_data}
PM_MB_means <- read.csv(here("cache/1911_PM_MB_means&Ygains.csv"))

source(here("R/same.R")) # matches each element of a vector and does not return NAs
```

@@

We are narrowing the meta-analysis to only include treatments using demethylation inhibitors, tebuconazole and propiconazole.
Let's look at what type of statistical error was reported for the trials that used these fungicides.

```{r variance_types}
PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(trial_ref, location, year , Y_error_type, fungicide_ai) %>%
   summarise(DMI_treatments = length(Y_error_type)) %>%
   arrange(Y_error_type) %>%
   select(!fungicide_ai) %>%
   kable(caption = "Number of treatments for each trial and the error type reported",
         align = "c") %>%
   footnote(general = "DMI - demethylation inhibitors (tebuconazole and propiconazole).")
```

The experiment summaries report either standard deviation, least square differences or no error term (`NA`). 
First we will convert LSD and standard deviation to sample variance.
Then we will impute variances for the trials, which did not report any form of variance (`NA`).

@@

First let's get a list the of trials that reported LSD.

```{r all_LSD_trials}
Trials <- PM_MB_means %>%
   filter(Y_error_type == "lsd (P=0.05)") %>%
   select(trial_ref) %>%
   distinct()
Trials
```

Let's use this list of trials to subset the data and calculate sampling variances.

We will follow the method of converting 'least square differences' (LSD) to sampling variance reported by [Nugugi et.al (2011)](https://apsjournals.apsnet.org/doi/10.1094/PHYTO-08-10-0221).
A [table of T-critical values](https://www.statisticshowto.datasciencecentral.com/tables/t-distribution-table/) was consulted to select the T-crit value in the equation. 
Our data contained 30 degrees of freedom which equates to a 'T-crit' value of 1.697.

@@

`Tcrit` is the T critical value for which there is significant difference between two groups.
This value is the same for both trials "mung1112/01" and "mung1112/02".

```{r LSD_2_Sampling_Variance}
Tcrit <- 2.042 # 0.05 and DFw = 30

for (i in Trials$trial_ref) {
   DFw <-
      sum(PM_MB_means[PM_MB_means$trial_ref == i, ]$replicates) -
      length(PM_MB_means[PM_MB_means$trial_ref == i, ]$replicates) # degrees of freedom within groups

   LSD <-
      PM_MB_means[PM_MB_means$trial_ref == i, ]$yield_error[1]

   V_yield <-
      (PM_MB_means[PM_MB_means$trial_ref == i, ]$replicates[1] * 
          (LSD / Tcrit) ^ 2) / 2 
   # method - Nugugi et.al (2011) DOI: 10.1094/phyto-08-10-0221

   for (j in seq_along(PM_MB_means[PM_MB_means$trial_ref == i,]$replicates)) {
      if (j == 1) {
         Sv <-
            vector(length = length(PM_MB_means[PM_MB_means$trial_ref == i,]$replicates))
      } # reset Sv vector if this is the first loop

      Sv[j] = V_yield /
         PM_MB_means[PM_MB_means$trial_ref == i,]$replicates[j]

@@

   # Replace the values in the dataset
   PM_MB_means[PM_MB_means$trial_ref == i,]$yield_error <-
      Sv
   PM_MB_means[PM_MB_means$trial_ref == i,]$Y_error_type <-
      "SamplingVariance"
}
```

### Converting standard deviation to sample variance

Now to estimate the variance from the standard deviation.

@@

# and replacing the values in the dataset
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$yield_error <-
   Sv
PM_MB_means[PM_MB_means$trial_ref %in% Trials$trial_ref,]$Y_error_type <-
   "SamplingVariance"
```

### Check variance of data-set
Our data currently uses sample variance to define the yield error.
We should expect that when yield increases so does the variance.
Let's quickly plot this to check.

```{r Var_mean_sqaures}
# redefine sample variance calculated from mean square
PM_MB_means$vi <-
   PM_MB_means$Y_Msquare /
   (PM_MB_means$n)

PM_MB_means %>%
   ggplot(aes(x = grain_yield.t.ha., y = vi)) +
   geom_point()
```

We see here that the sample variance generally decreases or has no association with the grain yield.
This might be due to the unpredictable nature of mungbeans, however going forward we will use mean squares, which defines between trial variation and is the more generic error term.

```{r Var_MeanSquarePlot}
PM_MB_means %>%
   ggplot(aes(x = grain_yield.t.ha., y = Y_Msquare)) +
   geom_point()
```

## Reduce data set to required variables

Let's tidy up the data set into a data frame of only the variables, which are required for the analysis.

First remove the irrelevant columns/variables.
Then calculate fungicide application timing variables relative to the first sign of disease.
Entries with `NA` in the grain yield or powdery mildew severity are also removed from the data.

```{r clustered_fungicide_applications}
source("R/slimming_PM_dat.R") # see R script for the code, which reduces the data.

slim_PM_dat %<>%
   mutate(fungicide_timing_1 = fungicide_application_1 - first_sign_disease) %>%
   mutate(fungicide_timing_2 = fungicide_application_2 - fungicide_application_1) %>%
   mutate(fungicide_timing_3 = fungicide_application_3 - fungicide_application_2) %>%
   filter(!is.na(grain_yield.t.ha)) %>%
   filter(!is.na(PM_final_severity))
```

### Cluster fungicide applications

We are going to define 'fungicide application timing' as a categorical variable, relative to the first sign of powdery mildew.

These categorical variables are named:

   - **Early**: First fungicide application was prior to first sign of disease.

   - **Recommended**: First fungicide application was applied on the day powdery mildew was observed, or within three days of first sign.

   - **Late**: First fungicide application was four or more days after first sign of disease being observed. 
Follow up fungicide applications (cluster_2 and cluster_3) were also categorised as either early, recommended or late.

```{r cluster}
slim_PM_dat %<>%
   mutate(
      cluster_1 = case_when(
         fungicide_timing_1 < 0 ~ "Early",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 ~ "Recommended",
         TRUE ~ "Late"
      )
   ) %>%
   mutate(
      cluster_2 = case_when(
         is.na(fungicide_timing_2) ~ NA_character_,
         fungicide_timing_2 < 13 ~ "Early",
         fungicide_timing_2 >= 13 &
            fungicide_timing_2 < 17 ~ "Recommended",
         TRUE ~ "Late"
      )
   ) %>%
   mutate(
      cluster_3 = case_when(
         fungicide_timing_3 < 14 ~ "Early",
         is.na(fungicide_timing_3) ~ NA_character_,
         TRUE ~ "Recommended"
      )
   )
```

Let's look at these clusters to see where we have the most degrees of freedom.

```{r cluster_clusters}
clusters <-
   unique(slim_PM_dat[c("cluster_1", "cluster_2", "cluster_3")])

for (i in 1:nrow(clusters)) {
   if (i == 1) {
      clusters$n <- NA
      clusters$n_x_trial <- NA
   }

   clusters[i, "n"] <-
      nrow(slim_PM_dat[same(slim_PM_dat$cluster_1, clusters[i, "cluster_1"]) &
                          same(slim_PM_dat$cluster_2, clusters[i, "cluster_2"]) &
                          same(slim_PM_dat$cluster_3, clusters[i, "cluster_3"]),])

   clusters[i, "n_x_trial"] <-
      length(unique(slim_PM_dat[same(slim_PM_dat$cluster_1, clusters[i, "cluster_1"]) &
                                   same(slim_PM_dat$cluster_2, clusters[i, "cluster_2"]) &
                                   same(slim_PM_dat$cluster_3, clusters[i, "cluster_3"]),
                                "trial_ref"]))

@@

knitr::kable(clusters)
```

From this we can see that there are 16 different cluster combinations, which would make the analysis too difficult. We will reduce this to 6 levels within one variable.

Single sprays of either:

   a) early,

   b) recommended timing,

   c) late.

Two or more sprays where the first spray was either:

   d) early,

   e) recommended timing,

   f) late.

```{r simple_clusters}
slim_PM_dat <- slim_PM_dat %>%
   mutate(
      spray_management = case_when(
         fungicide_timing_1 < 0 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Early",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Recommended",
         fungicide_timing_1 >= 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Late",
         fungicide_timing_1 < 0 &
            !is.na(fungicide_application_2) ~ "Early_plus",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            !is.na(fungicide_application_2) ~ "Recommended_plus",
         fungicide_timing_1 >= 4 &
            !is.na(fungicide_application_2) ~ "Late_plus",
         TRUE ~ "Other"
      )
   )

slim_PM_dat[slim_PM_dat$fungicide_ai == "control",
            c(
               "fungicide_timing_1",
               "fungicide_timing_2",
               "fungicide_timing_3",
               "spray_management"
            )] <- "control"
```

Now to view the number break-down of the `spray_management` treatments

```{r tableSprayManagement}
table(slim_PM_dat$spray_management)
```

There are few 'Early_plus' treatments, these treatments therefore will have too few comparisons with other treatments in the meta-analysis to provide accurate results. Therefore we will remove 'Early_plus' from the analysis.

```{r simpler_clusters_remove}
slimmer_PM_dat <-
   slim_PM_dat[slim_PM_dat$spray_management != "Early_plus", ]
```

****

## Test dose effect

Earlier we noted there were different doses of propiconazole, which might influence our analysis.
Let's check this by comparing a linear mixed effect model including and excluding dose.
We already have a factor breaking down the dose in the `trial` variable.
Let's make a new one without dose.

```{r dose_test}
slimmer_PM_dat$trial_noDose <- paste(slimmer_PM_dat$trial_ref, slimmer_PM_dat$year,
                 slimmer_PM_dat$location,slimmer_PM_dat$host_genotype,
                 slimmer_PM_dat$row_spacing,sep = "_")

cbind(Dose = head(slimmer_PM_dat$trial),
      NoDose = head(slimmer_PM_dat$trial_noDose)) %>%
   kable()
```

Let's test the dose effect in a basic linear mixed effect model.

```{r dose_lmer}
m1 <-
   lmer(log(grain_yield.t.ha * 1000) ~ factor(spray_management) +
           (1 | trial),
        data = slimmer_PM_dat)

m2 <-
   lmer(log(grain_yield.t.ha * 1000) ~ factor(spray_management) +
           (1 | trial_noDose),
        data = slimmer_PM_dat)

anova(m1, m2)
```

Using the model without dose produces a lower AIC, however we should accept the data with dose as there is a significant difference between the models and we should accept the more complicated model `m1`.
We will retain dose as a variable inside the `trial` factor.

```{r saveDat}
slimmer_PM_dat <-
   slimmer_PM_dat[, colnames(slimmer_PM_dat) != "trial_noDose"] #remove second trial column
```

## Disease pressure factor

We may wish to investigate the impact of spray management in mitigating the yield loss through the effect on the disease pressure.
We will set a categorical variable to indicate the amount of disease pressure in the trial, which will be evaluated on the control plot.
We will create two categories, `lowD` and `highD`.
We will use the median `AUDPC` to separate the trials into these two groups.

```{r DiseasePressure}
slimmer_PM_dat$D_pres <- NA_character_

Trial_Dpress <- slimmer_PM_dat %>%
   group_by(trial_ref) %>%
   filter(fungicide_ai == "control") %>%
   summarise(AUDPC_C = mean(AUDPC_m, na.rm = TRUE))

median_AUDPC <-
   median(as.vector(Trial_Dpress$AUDPC_C), na.rm = TRUE)
median_AUDPC

write.csv(median_AUDPC, here("cache/median_AUDPC_C.csv"), row.names = FALSE)

sort(Trial_Dpress$AUDPC_C)

for (i in unique(slimmer_PM_dat$trial_ref)) {
   if (is.na(Trial_Dpress[Trial_Dpress$trial_ref == i, "AUDPC_C"])) {
      next()
   }

   if (Trial_Dpress[Trial_Dpress$trial_ref == i, "AUDPC_C"] < median_AUDPC)
   {
      slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "D_pres"] <- "lowD"
   } else{
      if (Trial_Dpress[Trial_Dpress$trial_ref == i, "AUDPC_C"] >= median_AUDPC) {
         slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "D_pres"] <- "highD"
      } else{
         slimmer_PM_dat[slimmer_PM_dat$trial_ref == i, "D_pres"] <-
            NA_character_
      }
   }
}
```

Let's have a quick look to examine whether trials with high disease pressure sustain a greater yield loss.

```{r plot_disease_pressure}
slimmer_PM_dat %>%
   ggplot(aes(x = D_pres, y = grain_yield.t.ha))+
   geom_boxplot() +
   xlab("Disease pressure") +
   ylab("Grain yield (t/ha)")
```

From this plot we see that in the trials with high disease pressure tended to produce lower yields. 
There are `r length(unique(slimmer_PM_dat[is.na(slimmer_PM_dat$D_pres),"trial_ref"]))` trials that contained no data for AUDPC and thus produced `NA` values.

## Row spacing

We may want to consider row spacing in the trials.
However one of the locations has `NA` as the row spacing.
For now let's assign the most used row spacing of 0.33 m as an approximation so the model can analyse this as a continuous variable. 
If we decide to use the row_spacing column as a factor in the final model we will need to remove this approximation.

```{r assign_row_spacing_missing}
slimmer_PM_dat[is.na(slimmer_PM_dat$row_spacing),]
slimmer_PM_dat[is.na(slimmer_PM_dat$row_spacing),"row_spacing"] <- 0.33

write.csv(slimmer_PM_dat, file = here("cache/slimmer_PM_clusterdat.csv"), row.names = FALSE)
```

### Mungbeans age and yield saved

The time at which powdery mildew infects in the seasons and the age of the plant might affect the AUDPC and the loss in yield.
Let's have a look at the age of each trial when each powdery mildew infected it.
We will use the difference between `planting_date` and `first_sign_disease` as a proxy for crop maturity.

```{r age_andYield}
dat1 <- slimmer_PM_dat %>%
   mutate(days2PM = first_sign_disease - planting_date) %>%
   arrange(desc(days2PM)) %>%
   select(
      trial_ref,
      year,
      location,
      spray_management,
      days2PM,
      grain_yield.t.ha,
      yield_gain,
      prop_yield_gain,
      AUDPC_m,
      D_pres,
      PM_final_severity
   )
dat1
```

From these trials I am most interested in the experiments that directly compare `Recommended` with `Recommended_P_plus` and/or `Late` with `Late_plus`. 
Therefore I will not inspect, `AM1304` (Marys Mount).

We can see that the crops which were likely the most mature, did not show large yield gains, in particular:

   * Hermitage 2019  
   * Premer  
   * Goolhi  
   * Millmerran

Let's look at this in a quick box plot. 
We will only look at trials where powdery mildew set in at flowering which is approximately 60 days after planting.

```{r ageBoxPlot}
dat1 %>%
   filter(days2PM >= 60) %>%
   filter(spray_management != "control" &
             trial_ref != "AM1304") %>%
   group_by(trial_ref, spray_management) %>%
   ggplot(aes(x = spray_management, y = yield_gain, fill = trial_ref)) +
   geom_boxplot() +
   scale_fill_usq() +
   geom_hline(yintercept = 0)
```

We can see in these experiments five experiments there is clearly no benefit to applying a second fungicide application.

For completeness let's test this with powdery mildew infection commencing 40 - 60 days after planting, instead of greater than 60.

```{r PM40_2_59_days}
dat1 %>%
   filter(days2PM < 60 &
             days2PM >= 40) %>%
   filter(spray_management != "control") %>%
   group_by(trial_ref, spray_management) %>%
   ggplot(aes(x = spray_management, y = yield_gain, fill = trial_ref)) +
   geom_boxplot() +
   scale_fill_usq() +
   geom_hline(yintercept = 0)
```

We can see here all four experiments benefited, on average, from more than one spray.

Let's estimate the age of the crop based on crop maturity observations, we will use the `PM_MB_means` which include some observations of crop maturity.

```{r whatAreTheAges}
dat2 <- PM_MB_means %>%
   mutate(
      flower = lubridate::ymd(flowering_date) - lubridate::ymd(planting_date),
      pod = lubridate::ymd(pod_fill_date) - lubridate::ymd(planting_date),
      late_pod = lubridate::ymd(mid_late_pod_fill) - lubridate::ymd(planting_date)
   ) %>%
   select(location, year, flower, pod, late_pod) %>%
   distinct() %>%
   arrange(desc(flower))

index1 <-
   apply(dat2, 1 , function(X)
      ! all(is.na(X[c("flower", "pod", "late_pod")])))

dat2[index1,]
as.double(mean(dat2[index1, "flower"], na.rm = TRUE))
as.double(mean(dat2[index1, "pod"], na.rm = TRUE))
as.double(mean(dat2[index1, "late_pod"], na.rm = TRUE))
```

We can see that flowering and podding can be variable, and depend on the cumulative degree days.
However, on average flowering starts `r as.double(mean(dat2[index1,"flower"], na.rm = TRUE))` days after sowing, podding starts `r as.double(mean(dat2[index1,"pod"], na.rm = TRUE))`, and late podding `r as.double(mean(dat2[index1,"late_pod"], na.rm = TRUE))`.
